---
title: "2019-01-29- LCA data modeling Seth-Josh"
output: pdf_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE, message = FALSE, warning = FALSE, echo = TRUE)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60),  # For code
                      width = 60)  # For output

options(cli.width = 60)  # For tidyverse loading messages
set.seed(20180925)
knitr::clean_cache()
```

# 1. Loading, setting up

```{r, load}
library(tidyverse)
library(poLCA)
library(readxl)
```

# Getting data from Google Sheets

```{r, eval = TRUE}
library(googlesheets)
library(readr)

g <- gs_title("Observations_segment_Units_1-7_2013-14-with-duplicates-identified")
d <- gs_read(g, col_types = 
                 readr::cols(
                     `ClassObservation::Observer` = col_character(),
                     `ClassObservation::ObsNickname` = col_double(),
                     `Teacher::TeacherID` = col_double(),
                     `Teacher::First Name` = col_character(),
                     `Teacher::Last Name` = col_character(),
                     `Teacher::Condition` = col_character(),
                     `ClassObservation::Unit` = col_double(),
                     `ClassObservation::Date` = col_datetime(format = ""),
                     Notes = col_character(),
                     ObsNN = col_double(),
                     SegNum = col_double(),
                     `Segment::StartStamp` = col_datetime(format = ""),
                     `Segment::EndStamp` = col_datetime(format = ""),
                     fWhole = col_double(),
                     fGroups = col_double(),
                     fSeat = col_double(),
                     sInvented = col_double(),
                     sConceptual = col_double(),
                     sProcedural = col_double(),
                     sEngagement = col_character(),
                     tInitSelect = col_double(),
                     tCompare = col_double(),
                     tDiscussQ = col_double(),
                     tPressExplain = col_double(),
                     tConnectOthers = col_double(),
                     tConnectBigIdeas = col_double(),
                     tConventional = col_double(),
                     tProcedural = col_double(),
                     iPrecision = col_double(),
                     iCenter = col_double(),
                     iDIsplay = col_double(),
                     iOther = col_double(),
                     iOrder = col_double(),
                     iScale = col_double(),
                     iGrouping = col_double(),
                     iShape = col_double(),
                     iShow = col_double(),
                     iHide = col_double(),
                     iMode = col_double(),
                     iMedian = col_double(),
                     iMean = col_double(),
                     iRange = col_double(),
                     iCenterClump = col_double(),
                     iDeviation = col_double(),
                     iReplicability = col_double(),
                     iGeneralizability = col_double(),
                     iLinkVisDist = col_double(),
                     iLinkImagDist = col_double(),
                     ITheoreticalProb = col_double(),
                     IEmpiricalProb = col_double(),
                     IOdds = col_logical(),
                     ISampleSize = col_double(),
                     ISamplingDistrib = col_double(),
                     ICenterStats = col_double(),
                     IVariabilityStats = col_double(),
                     `Segment::iIntelligibility` = col_double(),
                     `Segment::iModelFit` = col_double(),
                     `Segment::iDistribution` = col_double(),
                     `Segment::iRandomComponents` = col_double(),
                     `Segment::iNonRandomComponents` = col_double(),
                     `Segment::iMedianDistr` = col_double(),
                     `Segment::iIQRDistr` = col_logical(),
                     `Segment::iNewMedian` = col_double(),
                     `Segment::iNewIQR` = col_logical(),
                     `Segment::iRegions` = col_double(),
                     `Segment::iQuantRegions` = col_double(),
                     number_of_segments = col_double(),
                     `Duplicate Condition` = col_character()
                 ))

d <- dplyr::rename(d, condition = `Teacher::Condition`)

d <- d %>% 
    mutate(condition = ifelse(str_detect(condition, "2"), 0,
                              ifelse(str_detect(condition, "1"), 1, NA)))

library(readxl)
u <- read_xlsx("Observations_summary_Units_1-7_2012-13-mod.xlsx")

g1 <- gs_title("Observations_segment_Units_1-7_2012-13-with-duplicates-identified")
d1 <- gs_read(g1)
d1 <- rename(d1, Teacher_ID = handl)
#d1 <- unite(d1, Teacher, `Teacher::First Name`, `Teacher::Last Name`, sep = " ")
d1 <- d1 %>% left_join(u, by = "Teacher_ID")
d1 <- rename(d1, condition = Group)
```


```{r}
add_one <- function(x) {
    x + 1
}

ds <- d %>% 
    dplyr::select(sInvented, sProcedural, sConceptual, tInitSelect, tCompare, tDiscussQ, tConnectBigIdeas, tConnectOthers, tPressExplain, condition, fGroups, fSeat, fWhole, `Duplicate Condition`, seg_num = SegNum, unit = `ClassObservation::Unit`, teacher = `Teacher::TeacherID`) %>% 
    map_df(replace_na, 0) %>%
    modify_at(c(1:9), add_one) %>% 
    mutate(groups = case_when(
        fGroups == 1 ~ "small_groups",
        fSeat == 1 ~ "seat",
        fWhole == 1 ~ "whole"
    )) %>% 
    dplyr::select(-fGroups, -fSeat, -fWhole)

ds1 <- d1 %>% 
    dplyr::select(sInvented, sProcedural, sConceptual, tInitSelect, tCompare, tDiscussQ, tConnectBigIdeas, tConnectOthers, tPressExplain, condition,  fGroups, fSeat, fWhole, `Duplicate Condition`, seg_num = SegNum, unit = `ClassObservation::Unit`, teacher = Teacher_ID) %>% 
    map_df(replace_na, 0) %>% 
    modify_at(c(1:9), add_one) %>% 
    mutate(groups = case_when(
        fGroups == 1 ~ "small_groups",
        fSeat == 1 ~ "seat",
        fWhole == 1 ~ "whole"
    )) %>% 
    dplyr::select(-fGroups, -fSeat, -fWhole)

dd <- bind_rows(ds, ds1)

dds <- filter(dd, `Duplicate Condition` != "D" & `Duplicate Condition` != "d")
```

# 3. Choosing the number of classes/profiles

Using latent class analysis through the **poLCA** R package.

```{r, mapping-fit-stats, results = 'hide', eval = TRUE}
set.seed(20180925)

f <- cbind(sInvented, sProcedural, sConceptual, tInitSelect, tCompare, tDiscussQ, tConnectBigIdeas, tConnectOthers, tPressExplain) ~ 1
#2341
od <- map(1:9, poLCA, formula = f, data = dds, maxiter = 5000, verbose = FALSE, graphs = FALSE) %>% 
    map_df(broom::glance)
```

```{r, plot, eval = TRUE}
od %>% 
    mutate(n_classes = 1:9) %>% 
    gather(key, val, BIC, AIC) %>% 
    ggplot(aes(x = n_classes, y = val, color = key, group = key)) +
    geom_point() +
    geom_line() +
    scale_x_continuous(breaks = 1:9, labels = 1:9) +
    theme_bw() +
    labs(caption = "Lower values of the AIC & BIC suggest preferred model(s); generally, BIC is more conservative than AIC") +
    xlab("Number of Classes") +
    ylab("Value") +
    scale_color_discrete("")
```

Based on these fit statistics, a three or four class solution seems to exhibit the best fit, though a three-class solution may also be suitable; for comparison, a two-class solution is also explored.

# 4. Examining 2, 3, 4, and 5 class solutions

```{r, spec-sols, cache = FALSE}
set.seed(20180925)

f <- cbind(sInvented, sProcedural, sConceptual, tInitSelect, tCompare, tDiscussQ, tConnectBigIdeas, tConnectOthers, tPressExplain) ~ 1

#m2 <- poLCA(f, dds, nclass = 2, maxiter = 10000, graphs = TRUE)
#m3 <- poLCA(f, dds, nclass = 3, maxiter = 10000, graphs = TRUE)
m4 <- poLCA(f, dds, nclass = 4, maxiter = 10000, graphs = TRUE)
#m5 <- poLCA(f, dds, nclass = 5, maxiter = 10000, graphs = TRUE)
```

```{r, eval = FALSE}
data.frame(dds, class = m4$predclass) %>% 
    dplyr::select(class, condition) %>% 
    count(class, condition)
```

# 5. Examining predictors of the 4-class solution - does not work well for 3 class solution

# Moving forward with four-class solution

```{r}
post_probs <- m4$posterior %>% as.data.frame() %>% setNames(paste0("C", 1:4, "_prob"))
df <- bind_cols(dds, post_probs)
df$class <- m4$predclass
df <- df %>% dplyr::select(-`Duplicate Condition`) %>% mutate_if(is.numeric, round, 3)   
write_csv(df, "2019-02-10-data-with-class-probs.csv")
```

# Plots


```{r, plot-func, include = FALSE}

f <- function(teacher_ID, condition, d) {
    
    ds <- d %>%
        mutate(tmp_val = 1) %>%
        mutate(unit = str_c("Unit ", unit)) %>%
        filter(teacher == teacher_ID)
    
    n_units <- length(unique(ds$unit))
    
    p <- ggplot(ds, aes(x = seg_num, y = tmp_val, color = class)) +
        geom_point() +
        theme_bw() +
        facet_wrap(~ unit, ncol = 5) +
        ggtitle(str_c("Four Class Solution for Teacher (ID): ", teacher_ID, " (", condition, ")")) +
        scale_y_continuous("", labels = NULL) +
        theme(axis.ticks.y = element_blank()) +
        xlab("Segment") +
        theme(legend.position= "top") +
        scale_color_discrete("") +
        labs(caption = "From Latent Class Analysis (LCA) using the poLCA R package")
    
    p 
    
    ggsave(str_c("plots-by-teacher/", teacher_ID, ".png"), plot = p, width = 10, height = 5)
    
    message(str_c("Sucessfully prepared plot for teacher (ID) ", teacher_ID))
}
```

```{r}
t <- df %>% 
    arrange(teacher, unit, seg_num) %>% 
    group_by(unit) %>% 
    summarize(max_seg_num = max(seg_num),
              max_unit = max(unit))

the_seqqer <- function(x) {
    seq(1, t$max_seg_num[x])
}

l <- list()
for (i in seq(t$max_seg_num)) {
    l[[i]] <- seq(1, t$max_seg_num[i])    
}

t$seq_l <- l

dtm <- dplyr::select(df, teacher, condition) %>% distinct() %>% arrange(teacher)

df$class <- as.factor(df$class)

df$class<- forcats::fct_recode(df$class, 
                               `Low Activity` = "1",
                               `Inventing & Connecting` = "2",
                               `Inventing & Discussing` = "3",
                               `Discussing Ideas` = "4")

teacher_ID = dtm$teacher
condition = dtm$condition
map2(teacher_ID, condition, f, df)
```

# Analysis

```{r}
dm1 <- df %>% count(class, condition) %>% 
    spread(condition, n) %>% 
    mutate(`0` = replace_na(`0`, 0))
names_dm1 <- dm1$class
mat1 <- as.matrix(dm1[, -1])
cs1 <- chisq.test(mat1)
write_csv(dm1, "tab1.csv")
write_csv(as.data.frame(cs1$stdres), "mat1.csv")
#clipr::write_clip(cs1$stdres)

dm2 <- df %>% count(class, groups) %>% 
    spread(groups, n)
names_dm2 <- dm2$class
mat2 <- as.matrix(dm2[, -c(1, 5)])
cs2 <- chisq.test(mat2)
write_csv(dm2, "tab2.csv")
write_csv(as.data.frame(cs2$stdres), "mat2.csv")
# clipr::write_clip(cs2$stdres)
```
