---
title: "LCA data modeling Seth-Josh"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE, message = FALSE, warning = FALSE, echo = TRUE)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60),  # For code
                      width = 60)  # For output

options(cli.width = 60)  # For tidyverse loading messages

set.seed(20180925)
```

# 1. Loading, setting up

```{r, load}
library(tidyverse)
library(poLCA)

f <- "obs-segment_units1-7_2013-2014.csv"

d <- read_csv(f)
```

# 2. Preparing data with a few teacher and student variables

None of the unit-specific variables included.

```{r, prep}
add_one <- function(x) {
    x + 1
}

ds <- d %>% 
    dplyr::select(sInvented, sProcedural, sConceptual, tInitSelect, tCompare, tDiscussQ, tConnectBigIdeas, tConnectOthers, tPressExplain, fGroups) %>% 
    map_df(replace_na, 0) %>% 
    map_df(add_one)
```

# 2. testing tidyLPA

```{r}
library(tidyLPA)
dss <- d %>% 
    dplyr::select(sInvented, sProcedural, sConceptual, tInitSelect, tCompare, tDiscussQ, tConnectBigIdeas, tConnectOthers, tPressExplain, fGroups) %>% 
    map_df(replace_na, 0)
estimate_profiles_mplus(dss, sInvented, sProcedural, sConceptual, n_profiles=2, categorical_indicators=TRUE, remove_tmp_files=FALSE, variances = "equal", covariances = "equal")
```

# 3. Choosing the number of classes/profiles

Using latent class analysis through the **poLCA** R package.

```{r, mapping-fit-stats, results = 'hide', cache = TRUE, eval = FALSE}
f <- cbind(sInvented, sProcedural, sConceptual, tInitSelect, tCompare, tDiscussQ, tConnectBigIdeas, tConnectOthers, tPressExplain) ~ 1

f <- cbind(sInvented, sProcedural, sConceptual, tPressExplain) ~ 1

od <- map(1:9, poLCA, formula = f, data = ds, maxiter = 5000, verbose = FALSE, graphs = FALSE) %>% 
    map_df(broom::glance)

saveRDS(od, "2013-2014-dat-compare.rds")
```

```{r}
od <- read_rds("2013-2014-dat-compare.rds")
```

```{r, plot}
od %>% 
    mutate(n_classes = 1:9) %>% 
    gather(key, val, BIC, AIC) %>% 
    ggplot(aes(x = n_classes, y = val, color = key, group = key)) +
    geom_point() +
    geom_line() +
    scale_x_continuous(breaks = 1:9, labels = 1:9) +
    theme_bw() +
    labs(caption = "Lower values of the AIC & BIC suggest preferred model(s); generally, BIC is more conservative than AIC")
```

Based on this fit statistic--the Bayesian Information Criteria, which is just a transformation of the log-likelihood, and is usually recommended along with the AIC as one criterion for model selection--it looks like 3 and especially 4 or 5 class solutions seem reasonable.

# 4. Examining 3, 4, and 5 class solutions

```{r, spec-sols}
f <- cbind(sInvented, sProcedural, sConceptual, tInitSelect, tCompare, tDiscussQ, tConnectBigIdeas, tConnectOthers, tPressExplain) ~ 1

m3 <- poLCA(f, ds, nclass = 3, maxiter = 5000, graphs = TRUE)
m4 <- poLCA(f, ds, nclass = 4, maxiter = 5000, graphs = TRUE)
m5 <- poLCA(f, ds, nclass = 5, maxiter = 5000, graphs = TRUE)
```

# 5. Examining the effects of a covariate (fGroups)

```{r, cov}
fcov <- cbind(sInvented, sProcedural, sConceptual, tInitSelect, tCompare, tDiscussQ, tConnectBigIdeas, tConnectOthers, tPressExplain) ~ fGroups

m4cov <- poLCA(fcov, ds, nclass = 4, maxiter = 5000, graphs = TRUE)
```

- What covariates do we want to add?

# 6. Plot of segments over time for the four class solution

```{r, plot-func, include = FALSE}

f <- function(teacher_ID, condition, d) {
    
    ds <- d %>%
        dplyr::select(seg_num = SegNum, unit = `ClassObservation::Unit`, teacher = `Teacher::TeacherID`, m4_class_name, condition = `Teacher::Condition`) %>% 
        mutate(tmp_val = 1) %>%
        mutate(unit = str_c("Unit ", unit)) %>%
        filter(teacher == teacher_ID)
    
    n_units <- length(unique(ds$unit))
    
    p <- ggplot(ds, aes(x = seg_num, y = tmp_val, color = as.factor(m4_class_name))) +
        geom_point() +
        theme_bw() +
        facet_wrap(~ unit, ncol = 5) +
        ggtitle(str_c("Four Class Solution for Teacher (ID): ", teacher_ID, " (", condition, ")")) +
        scale_y_continuous("", labels = NULL) +
        theme(axis.ticks.y = element_blank()) +
        xlab("Segment") +
        theme(legend.position= "top") +
        scale_color_discrete("") +
        labs(caption = "From Latent Class Analysis (LCA) using the poLCA R package")
    
    p 
    
    ggsave(str_c("plots-by-teacher/", teacher_ID, ".png"), plot = p, width = 10, height = 5)
    
    message(str_c("Sucessfully prepared plot for teacher (ID) ", teacher_ID))
    
}

```

```{r, results = 'hide', four-class}
d$m4_class <- m4$predclass

t <- d %>% 
    dplyr::select(seg_num = SegNum, unit = `ClassObservation::Unit`, teacher = `Teacher::TeacherID`) %>% 
    arrange(teacher, unit, seg_num) %>% 
    group_by(unit) %>% 
    summarize(max_seg_num = max(seg_num),
              max_unit = max(unit)) 

the_seqqer <- function(x) {
    seq(1, t$max_seg_num[x])
}

l <- list()
for (i in seq(t$max_seg_num)) {
    l[[i]] <- seq(1, t$max_seg_num[i])    
}

t$seq_l <- l

dtm <- dplyr::select(d, `Teacher::TeacherID`, `Teacher::Condition`) %>% distinct() %>% arrange(`Teacher::TeacherID`)

d$m4_class <- as.factor(d$m4_class)

levels = c("Discussing Ideas", "Inventing & Discussing", "Inventing & Connecting", "Low Activity")

d$m4_class_name <- forcats::fct_recode(d$m4_class, 
                                       `Discussing Ideas` = "1",
                                       `Inventing & Connecting` = "3",
                                       `Inventing & Discussing` = "2",
                                       `Low Activity` = "4")

```

```{r}
teacher_ID = dtm$`Teacher::TeacherID`
condition = dtm$`Teacher::Condition`
map2(teacher_ID, condition, f, d)
```

# 7. Adding extra data

```{r}
post_probs <- m4$posterior %>% as.data.frame() %>% setNames(paste0("C", 1:4, "_prob"))
ds <- bind_cols(d, post_probs)
ds <- rename(ds, class = m4_class)
write_csv(ds, "data-with-class-probs.csv")
```

# 8. Other ideas

- Examine unit-specific indicators by class?
- Examine control vs. intervention mean proportions?

# 9. Examining homophily and separation

```{r}

```
